---
title: "R Script for Data Analytics in Finance using R - Project"
date: "2022-04-30"
output: html_document
---

#########################################################################
<b>Initialize & import required libraries</b>
```{r message=FALSE}
rm(list=ls())
if("dplyr" %in% rownames(installed.packages()) == FALSE)
  install.packages("dplyr")
library(dplyr)
if("tidyr" %in% rownames(installed.packages()) == FALSE)
  install.packages("tidyr")
library(tidyr)
if("ggplot2" %in% rownames(installed.packages()) == FALSE)
  install.packages("ggplot2")
library(ggplot2)
if("gridExtra" %in% rownames(installed.packages()) == FALSE)
  install.packages("gridExtra")
library(gridExtra)
```


<b>Import dataset (master)</b>
Data source: https://github.com/owid/covid-19-data

```{r data import}
ds <- read.csv("owid-covid-data.csv")
```


<b>Normalize and cleaning</b>


Normalize Date/time
```{r normalize}
ds$date <- as.Date(ds$date, "%Y-%m-%d");
```


Filter location data and statistic data
```{r filter location}
dataLocation <- ds %>%
  filter(!grepl('^OWID.*',iso_code)) %>% # Remove other data not related to continent
  filter(date >= '2020-03-01' & date <= '2022-03-31') # Use data start from March20 (Since Feb20's data is incomplete)
dataStat <- ds %>%
  filter(grepl('^OWID.*',iso_code)) %>% # OWID_* is statistic data from OurWorld, we will not use this data. 
  filter(date >= '2020-03-01' & date <= '2022-03-31')
```


<b>Part 1: Visualize the Data</b>


Slide 5 - 7: Continent Level Daily Trend - new cases / new deaths / mortality rate
```{r slide 5-7, warning=FALSE,message=FALSE}
# Aggregate continent level data
dataContinent <- ds %>%
  group_by(continent, date) %>%
  summarize(continental_new_cases=sum(new_cases_smoothed,na.rm = T),
            continental_new_deaths=sum(new_deaths_smoothed,na.rm = T),
            continental_new_vaccinations=sum(new_people_vaccinated_smoothed,na.rm = T),
            continental_total_cases=sum(total_cases,na.rm=T),
            continental_total_deaths=sum(total_deaths,na.rm=T)) %>%
  mutate(continental_mortality_rate=continental_total_deaths/continental_total_cases)

#plot line chart to show the trend of new cases, new deaths and mortality per continent
ggplot(data = dataContinent, 
       aes(x=date, y=continental_new_cases, group=continent)) + 
  geom_line(aes(color=continent))+
  labs(title="Daily New Cases Trend by Continent",x="date",y="new_cases")

ggplot(data = dataContinent, 
       aes(x=date, y=continental_new_deaths, group=continent)) + 
  geom_line(aes(color=continent))+
  labs(title="Daily New Deaths Trend by Continent",x="date",y="new_death")

ggplot(data=dataContinent,
       aes(x=date, y=continental_mortality_rate,group=continent)) + 
  geom_line(aes(color=continent))+
  labs(title="Mortality Rate Trend by Continent",x="date",y="mortality_rate")
```


Slide 8-9: Global Distribution of total Deaths and Vaccination
```{r slide 8-9, warning=FALSE}

#Count total death by Continent / Country
vacc_ct <- dataLocation %>%
  group_by(continent) %>%
  summarize(mydata = sum(new_vaccinations_smoothed, na.rm=T))
death_ct <- dataLocation %>%
  group_by(continent) %>%
  summarize(mydata = sum(new_deaths_smoothed, na.rm=T))

#Plot graphs
g1 <- ggplot(data = death_ct) + 
  geom_col(aes(x=continent, y=mydata, fill=continent)) +
  ggtitle("No. of deaths by Continent") +
  xlab("Continent") + ylab("No. of people") +
  theme(plot.title = element_text(hjust = 0.5))

g2 <- ggplot(data = vacc_ct) + 
  geom_col(aes(x=continent, y=mydata, fill=continent)) +
  ggtitle("No. of Vaccinated people by Continent") +
  xlab("Continent") + ylab("No. of people") +
  theme(plot.title = element_text(hjust = 0.5))
g1
g2
```


Slide 10-11: Global Map
```{r slide 10-11, warning=FALSE}
#Graph the world map, import required libraries
if("maps" %in% rownames(installed.packages()) == FALSE)
  install.packages("maps")
library(maps)
if("RColorBrewer" %in% rownames(installed.packages()) == FALSE)
  install.packages("RColorBrewer")
library(RColorBrewer)
if("forcats" %in% rownames(installed.packages()) == FALSE)
  install.packages("forcats")
library(forcats)

#Create world map
world_map <- map_data(map = "world")
world_map$region <- iso.alpha(world_map$region, n=3) # Convert region code to ISO code 3 digit

#Prepare data - Total deaths percentage by location
df <- dataLocation %>%
  select(iso_code, total_deaths, population) %>%
  filter(total_deaths >= 0) %>%
  group_by(iso_code) %>%
  summarize(data = max(100*(total_deaths/population), na.rm=T))

#Define map scales
df$data_scale <- cut(df$data,
                      breaks = c(-Inf, 0.005, 0.01, 0.05, 0.1, 0.3, 0.5, Inf), 
                      labels = c("Less than 0.005%", "0.01%-0.005%", "0.05%-0.01%", 
                                 "0.1%-0.05%", "0.3%-0.1%", "0.5-0.3%", "More than 0.5%"))
#Plot the map of death %, with 7 level scales in red, country border in black and do not fill area
ggplot(df) +
  geom_map(aes(map_id = iso_code, fill = fct_rev(data_scale)), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'black', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  scale_fill_manual(name = "Total Deaths %", values = rev(brewer.pal(7, name = "Reds"))) +
  theme_void() +
  coord_fixed()

#Vaccinated people % by location
df <- dataLocation %>%
  select(iso_code, people_vaccinated, population) %>%
  filter(people_vaccinated >= 0) %>%
  group_by(iso_code) %>%
  summarize(data = max(people_vaccinated/population, na.rm=T))

df$data_scale <- cut(df$data, 
                     breaks = c(-Inf, 0.05, 0.1, 0.2, 0.4, 0.6, 0.9, Inf), 
                     labels = c("Less than 5%", "10%-5%", "20%-10%", "40%-20%", 
                                "60%-40%", "90%-60%", "More than 90%"))

world_map <- map_data(map = "world")
world_map$region <- iso.alpha(world_map$region, n=3)

ggplot(df) +
  geom_map(aes(map_id = iso_code, fill = fct_rev(data_scale)), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'black', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  scale_fill_manual(name = "Vaccination %", values = rev(brewer.pal(7, name = "Greens"))) +
  theme_void() +
  coord_fixed()
```


Slide 12: Global Distribution of total cases
```{r slide 12, warning=FALSE}
#Pie chart of total cases as at 2020-12-31 by continents
ds_year_end_20<-ds %>% 
  filter(date=="2020-12-31") %>% 
  filter(location %in% c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
  ) %>%
  select("continent", "location", "date", "total_cases") %>% 
  mutate(percentage_total_cases=round(total_cases/sum(total_cases)*100, digits=0))

data_Pie_2020<-data.frame("category"=ds_year_end_20$location, 
                          "amount"=ds_year_end_20$percentage_total_cases)

Pie_A<-ggplot(data_Pie_2020, aes(x="", y=amount, fill=category)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  geom_text(aes(label = paste0(amount,"%" )), position = position_stack(vjust=0.5)) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  ggtitle('Proportion of total cases in different continents in 2020')

#Pie chart of total cases as at 2021-12-31 by continents
ds_year_end_21<-ds %>% 
  filter(date=="2021-12-31") %>% 
  filter(location %in% c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
  ) %>% 
  select("continent", "location", "date", "total_cases") %>% 
  mutate(percentage_total_cases=round(total_cases/sum(total_cases)*100, digits=0))

data_Pie_2021<-data.frame("category"=ds_year_end_21$location, 
                          "amount"=ds_year_end_21$percentage_total_cases)

Pie_b<-ggplot(data_Pie_2021, aes(x="", y=amount, fill=category)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  geom_text(aes(label = paste0(amount,"%" )), position = position_stack(vjust=0.5)) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  ggtitle('Proportion of total cases in different continents in 2021')

#Pie chart of total cases as at 2022-3-31 by continents, being the cut off date of data
ds_period_end_22<-ds %>% 
  filter(date=="2022-3-31") %>% 
  filter(location %in% c("Africa", "Asia", "Europe", "North America", "Oceania", "South America")
  ) %>% 
  select("continent", "location", "date", "total_cases") %>% 
  mutate(percentage_total_cases=round(total_cases/sum(total_cases)*100, digits=0))

data_Pie_2022<-data.frame("category"=ds_period_end_22$location, 
                          "amount"=ds_period_end_22$percentage_total_cases)

Pie_c<-ggplot(data_Pie_2022, aes(x="", y=amount, fill=category)) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  geom_text(aes(label = paste0(amount,"%" )), position = position_stack(vjust=0.5)) +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme_classic() +
  theme(axis.line = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())+
  ggtitle('Proportion of total cases in different continents in 2022')

grid.arrange(Pie_A,Pie_b,Pie_c, nrow=1)
```


Slide 13:Asia Data
```{r slide 13, warning = FALSE}
#Prepare Data
ct <- "Asia"
vacc_loc <- dataLocation %>%
  filter(continent == ct) %>%
  group_by(location) %>%
  summarize(mydata = sum(new_vaccinations_smoothed, na.rm=T))
death_loc <- dataLocation %>%
  filter(continent == ct) %>%
  group_by(location) %>%
  summarize(mydata = sum(new_deaths_smoothed, na.rm=T))

#Plot the graphs
g1 <- ggplot(data = death_loc) +
  geom_col(aes(x=location, y=mydata, fill=location)) + 
  ggtitle("No. of deaths in Asia") +
  xlab("Location") + ylab("No. of people") +
  #theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  theme(legend.position="none")

g2 <- ggplot(data = vacc_loc) +
  geom_col(aes(x=location, y=mydata, fill=location)) + 
  ggtitle("No. of vaccinated people in Asia") +
  xlab("Location") + ylab("No. of people") +
  #theme(plot.title = element_text(hjust = 0.5)) +
  coord_flip() +
  theme(legend.position="none")

grid.arrange(g1, g2, nrow=1) # One screen two plots
```


Slide 14 - 15: Focus on Hong Kong
```{r slide 14-15, warning=FALSE}
#Prepare data

dataHK <- dataLocation %>%
  filter(iso_code == "HKG")
#Plot the number of death vs cases against by date (from a specific date range)

#Select values for plotting graph
dataHK.p <- dataHK %>% 
  select(date, new_cases_smoothed, new_tests_smoothed)
# dataHK.p <- dataHK.p[complete.cases(dataHK.p),] # Select non-NA rows only
dataHK.p <- na.omit(dataHK.p)

# View(dataHK.p)
ggplot(data = dataHK.p) + 
  geom_line(aes(x=date, y=new_tests_smoothed), colour="red") +
  geom_line(aes(x=date, y=new_cases_smoothed), colour="blue") +
  scale_x_date(limit=c(as.Date("2021-03-01"), as.Date("2022-03-31"))) +
  ggtitle("Case and death against date in HK") +
  xlab("Date") + ylab("COVID19 Cases and Death") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values = c("red", "blue"))

#Graph the total cases and vaccinations, preparing data
dataHK.p2 <- dataHK %>%
  select(date, total_cases, total_vaccinations, new_tests_smoothed, total_deaths)

# nrow(dataHK.p2)
# View(dataHK.p2)
# Plot some simple graphs about HK (cases/vaccinations/tests/deaths)
g1 <- ggplot(data = dataHK.p2) +
  geom_line(aes(x=date, y=total_cases), colour="red") + labs(title="cases vs date")
g2 <- ggplot(data = dataHK.p2) +
  geom_line(aes(x=date, y=total_vaccinations), colour="blue") + labs(title="vaccinations vs date")
g3 <- ggplot(data = dataHK.p2) +
  geom_line(aes(x=date, y=new_tests_smoothed), colour="black")+ labs(title="new tests vs date")
g4 <- ggplot(data = dataHK.p2) +
  geom_line(aes(x=date, y=total_deaths), colour="green") + labs(title="total deaths vs date")
grid.arrange(g1, g2, g3, g4, nrow=2, ncol=2)
```


Slide 16 - 17: Comparing HK with Other Developed Countries in Asia
```{r slide 16-17, warning= FALSE}
#Prepare Data
loc <- c("HKG", "TWN", "SGP", "KOR", "JPN", "MYS")
analysis <- c("new_cases_smoothed_per_million", "new_vaccinations_smoothed_per_million", 
              "new_tests_smoothed_per_thousand", "new_deaths_smoothed_per_million")
dataSel <- dataLocation %>%
  select(iso_code, location, date, any_of(analysis)) %>%
  filter(iso_code %in% loc)

dataSel <- na.omit(dataSel) # Remove NA

#Plot the graphs
g1 <- ggplot() +
  geom_line(data = filter(dataSel, iso_code == "HKG"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "TWN"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "SGP"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "KOR"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "JPN"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "MYS"), aes(x=date, y=new_cases_smoothed_per_million, group=iso_code, colour=iso_code)) +
  labs(title="New cases against date")
g2 <- ggplot() +
  geom_line(data = filter(dataSel, iso_code == "HKG"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "TWN"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "SGP"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "KOR"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "JPN"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "MYS"), aes(x=date, y=new_vaccinations_smoothed_per_million, group=iso_code, colour=iso_code)) +
  labs(title="New vaccinations against date")
g3 <- ggplot() +
  geom_line(data = filter(dataSel, iso_code == "HKG"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "TWN"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "SGP"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "KOR"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "JPN"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "MYS"), aes(x=date, y=new_tests_smoothed_per_thousand, group=iso_code, colour=iso_code)) +
  labs(title="New tests against date")
g4 <- ggplot() +
  geom_line(data = filter(dataSel, iso_code == "HKG"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "TWN"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "SGP"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "KOR"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "JPN"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  geom_line(data = filter(dataSel, iso_code == "MYS"), aes(x=date, y=new_deaths_smoothed_per_million, group=iso_code, colour=iso_code)) +
  labs(title="New deaths against date")

grid.arrange(g1, g4, nrow=2)
grid.arrange(g3, g2, nrow=2)
```


Slide 18: Mortality Rate Comparison
```{r slide 18, warning=FALSE}
# Get Constant stats for each location
dataLocationCons <- ds %>%
  filter(date=="2021-12-31") %>%
  select(contains("per"),population_density) %>%
  select(-starts_with("new"))%>%
  select(-starts_with("weekly")) %>%
  mutate(mortality_rate=total_deaths_per_million/total_cases_per_million)
newdf<-dataLocationCons[!is.na(dataLocationCons$mortality_rate),]
# View(newdf)

# Location Mortality Rate
dataLocMR <- ds %>%
  filter(date=="2022-04-15")%>%
  mutate(mortalityrate=total_deaths_per_million/total_cases_per_million)%>%
  filter(!mortalityrate==0) %>%
  mutate(meanMR=mean(mortalityrate))
View(dataLocMR)

g<-ggplot(data=dataLocMR,
       mapping=aes(x=location,y=mortalityrate,color=continent))+
  geom_col()+
  coord_flip()+
  theme(axis.text.x = element_text(size = 6, vjust = 0.5, hjust = 0.5, angle = 45))+
  theme(axis.text.y = element_text(size = 2, vjust = 0.5, hjust = 0.5))

g+geom_point(data=dataLocMR,
            mapping=aes(x=location, y=meanMR,color="red",))+
  theme(legend.position = "none")
```


<b>Part 2: Correlation and regression analysis</b>


Slide 19: Herd Immunity
```{r slide 19, warning=FALSE}
# Get vaccination stats for each location
dataLocationVac <- dataLocation %>%
  select(iso_code,location,date,total_cases,total_deaths,
         continent,contains("vaccinations") |contains( "vaccinated"), contains("booster"))%>%
  mutate(mortality_rate=total_deaths/total_cases)

dataLocationVacN<-na.omit(dataLocationVac) #remove the locations not reporting vaccination data

LocListVac <- dataLocationVacN %>%
  group_by(location)%>%
  summarize(num=n())%>%
  filter(num>=300)%>%
  select(location) #to focus on countries reporting vaccination data more consistently

LocListVac <- unlist(LocListVac) #change list to character

# Mutate new variable to reflect herd immunity  
df2<- ds %>%
  filter(location %in% LocListVac) %>%
  select(location,date,total_cases_per_million) %>%
  rename(total_cases_per_million_T7=total_cases_per_million) # create a temp dataframe to save T-7 cases stat

dataLocationVacN2 <- dataLocationVacN %>%
  mutate(T7=date-7)

dataLocationHI <- left_join(dataLocationVacN2,df2,by=c('location'='location','T7'='date')) #add T-7 cases number to the original df
dataLocationHI <- dataLocationHI %>%
  mutate(HI=total_cases_per_million_T7/10000+people_vaccinated_per_hundred)%>%
  filter(location %in% LocListVac)

g31<- ggplot(dataLocationHI,
             aes(x=HI, y=mortality_rate, group=location))+
  geom_point(aes(color=date))+
  facet_wrap(~location)
g31
```


Slide 20: Elder Population % - defer to Earnest
```{r slide 20, message=FALSE, warning = FALSE}
#Bubble plot of average death rates vs percentage of elderly
daily_death_rate<-dataLocation %>% 
  select("continent", "location", "date", "total_cases", "total_deaths", "people_vaccinated_per_hundred", "population", "population_density", "aged_65_older") %>% 
  mutate(death_rate=total_deaths/total_cases*100)

avg_death_rate_location<-daily_death_rate %>% 
  group_by(location, continent) %>% 
  summarise(`Average death rates (%)`=round(mean(death_rate, na.rm = TRUE), digits=3), population=mean(population), `Percentage of population with age 65 or above (%)`=mean(aged_65_older)) 

ggplot(avg_death_rate_location, aes(x=`Percentage of population with age 65 or above (%)`, y=`Average death rates (%)`, size = population, color=continent)) +
  geom_point(alpha=0.5)
```


Slide 21-23: Examine the relation between diff factors and mortality rate
```{r slide 21-23, message=FALSE, warning = FALSE}
library(naniar)

factors_missing <- dataLocMR %>%
  miss_var_summary() # to identify amount of missing data for each factor

factors_not_missing <- factors_missing %>%
  filter(pct_miss<=50) %>%
  select(variable)

indexfactors<-c("location","total_cases_per_million","mortalityrate","continent")
econfactors <- c("human_development_index","extreme_poverty","gdp_per_capita")
hcfactors<-c("hospital_beds_per_thousand","diabetes_prevalence","cardiovasc_death_rate" )
demofactors <- c("median_age","population_density","life_expectancy")

# test economic factors
dataEconTest <- dataLocMR%>%
  select(econfactors,indexfactors)

g11 <- ggplot(data=dataEconTest,
              mapping=aes(x=human_development_index,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none")

g12 <- ggplot(data=dataEconTest,
              mapping=aes(x=gdp_per_capita,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none",axis.title.y=element_blank())

g13<- ggplot(data=dataEconTest,
             mapping=aes(x=extreme_poverty,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(axis.title.y=element_blank())

grid.arrange(g11,g12,g13,nrow=1)


# test healthcare factors
dataHCTest <- dataLocMR%>%
  select(hcfactors,indexfactors)

g21 <- ggplot(data=dataHCTest,
              mapping=aes(x=hospital_beds_per_thousand,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none")

g22 <- ggplot(data=dataHCTest,
              mapping=aes(x=diabetes_prevalence,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none",axis.title.y=element_blank())

g23 <- ggplot(data=dataHCTest,
              mapping=aes(x=cardiovasc_death_rate,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(axis.title.y=element_blank())

grid.arrange(g21,g22,g23,nrow=1)

# test demographic factors
dataDemoTest <- dataLocMR%>%
  select(demofactors,indexfactors)

g31 <- ggplot(data=dataDemoTest,
              mapping=aes(x=median_age,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none")

g32 <- ggplot(data=dataDemoTest,
              mapping=aes(x=population_density,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(legend.position = "none",axis.title.y=element_blank())

g33 <- ggplot(data=dataDemoTest,
              mapping=aes(x=life_expectancy,y=mortalityrate))+
  geom_point(aes(color=continent))+
  geom_smooth(na.rm=T)+
  theme(axis.title.y=element_blank())

grid.arrange(g31,g32,g33,nrow=1)
```


Slide 24: Linear Regression
```{r slide 24, warning = FALSE}
fit <- lm(mortalityrate ~ human_development_index+gdp_per_capita+hospital_beds_per_thousand+median_age+life_expectancy,data=dataLocMR)

summary(fit)
```


Slide 25: Decision Tree on Mortality Rate
```{r slide 25, warning = FALSE, message=FALSE}
#Import required libraries
if("caret" %in% rownames(installed.packages()) == FALSE)
  install.packages("caret")
library(caret)
if("rpart" %in% rownames(installed.packages()) == FALSE)
  install.packages("rpart")
library(rpart)

# Library for plot beautiful decision tree
if("rattle" %in% rownames(installed.packages()) == FALSE)
  install.packages("rattle")
library(rattle)

# Prepare Data
df <- ds %>%
  filter(!grepl('^OWID.*',iso_code))  %>% 
  filter(date >= '2020-03-01' & date <= '2022-03-31')
df <- df %>%
  select(new_deaths_smoothed_per_million, stringency_index, icu_patients_per_million, hosp_patients_per_million, weekly_icu_admissions_per_million,
         weekly_hosp_admissions_per_million, population_density, median_age, aged_65_older, aged_70_older, gdp_per_capita, 
         extreme_poverty, cardiovasc_death_rate, diabetes_prevalence, female_smokers, male_smokers, handwashing_facilities,
         hospital_beds_per_thousand, human_development_index)
# df <- do.call(data.frame, lapply(df, function(x) replace(x, is.infinite(x), NA)))
# View(df)
# na.omit(df)
# df[ , colSums(is.na(df)) == 0]
# Factor the death rate by mean + 1 std dev.
factor_mean <- mean(df$new_deaths_smoothed_per_million, na.rm=T) # + sd(df$new_deaths_smoothed_per_million, na.rm=T)

df$is_high <- ifelse(df$new_deaths_smoothed_per_million > factor_mean, 1, 0) # 1 = high death rate, 0 = low death rate
df <- df %>% drop_na(is_high)
df$is_high <- factor(df$is_high)
#
df <- df[-1] # Remove source factor column

set.seed(10)  # fix the randomizer
inTraining <- createDataPartition(df$is_high, p = 0.75, list = FALSE) 

training <- df[inTraining,] #75% data for model training  
testing <- df[-inTraining,] #25% for model testing

# nrow(df); nrow(training); nrow(testing)

fit <- rpart(is_high ~ ., data=training, method="class", 
             parms=list(split="information"))
# summary(fit)

pred <- predict(fit, type="class", newdata=testing)   # Predict Class
ConfusionMatrix <- table(pred, testing$is_high)      # Confusion Matrix Table
ConfusionMatrix
sum(diag(ConfusionMatrix))/sum(ConfusionMatrix)       # calculate the accuracy

#Plot the decision tree, black line, yes/no decision
fancyRpartPlot(fit, yesno=2, split.col="black", nn.col="black", 
               caption="", palette="Set2", branch.col="black")
```


Slide 26-29: Examine the relation between diff factors and new cases rate
```{r slide 26-29,warning=FALSE}
#Prepare new data, customize the date range
ds1 <- ds %>%
  filter(!grepl('^OWID.*',iso_code)) %>% 
  filter(date >= '2020-03-01' & date <= '2022-03-31')

# Selected factors to analysis
ds1 <- ds1 %>% 
  select(location, total_cases_per_million, stringency_index, icu_patients_per_million, hosp_patients_per_million, weekly_icu_admissions_per_million,
         weekly_hosp_admissions_per_million, population_density, median_age, aged_65_older, aged_70_older, gdp_per_capita, 
         extreme_poverty, cardiovasc_death_rate, diabetes_prevalence, female_smokers, male_smokers, handwashing_facilities,
         hospital_beds_per_thousand, human_development_index) %>%
  group_by(location) %>%
  summarise_if(is.numeric, max, na.rm=T) # One location one record, use max as grouped data
# View(ds1)
ds1 <- do.call(data.frame, lapply(ds1, function(x) replace(x, is.infinite(x), NA))) # Replace infinite value to NA
# class(ds1)


# Graph the correlation on each factors compared to total_cases_per_million
par(mfrow=c(2,2)) # 2 by 2 in one plot
plot(ds1$stringency_index, ds1$total_cases_per_million)
plot(ds1$human_development_index, ds1$total_cases_per_million)
plot(ds1$gdp_per_capita, ds1$total_cases_per_million)
plot(ds1$extreme_poverty, ds1$total_cases_per_million)
#
par(mfrow=c(3,2))
plot(ds1$icu_patients_per_million, ds1$total_cases_per_million)
plot(ds1$hosp_patients_per_million, ds1$total_cases_per_million)
plot(ds1$weekly_icu_admissions_per_million, ds1$total_cases_per_million)
plot(ds1$weekly_hosp_admissions_per_million, ds1$total_cases_per_million)
plot(ds1$hospital_beds_per_thousand, ds1$total_cases_per_million)
#
par(mfrow=c(2,2))
plot(ds1$population_density, ds1$total_cases_per_million)
plot(ds1$median_age, ds1$total_cases_per_million)
plot(ds1$aged_65_older, ds1$total_cases_per_million)
plot(ds1$aged_70_older, ds1$total_cases_per_million)
#
par(mfrow=c(3,2))
plot(ds1$cardiovasc_death_rate, ds1$total_cases_per_million)
plot(ds1$diabetes_prevalence, ds1$total_cases_per_million)
plot(ds1$female_smokers, ds1$total_cases_per_million)
plot(ds1$male_smokers, ds1$total_cases_per_million)
plot(ds1$handwashing_facilities, ds1$total_cases_per_million)

#Reset plot
par(mfrow=c(1,1))
```


Slide 30: Correlation matrix
```{r slide 30, warning = FALSE}
cor(ds1$stringency_index, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$human_development_index, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$gdp_per_capita, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$extreme_poverty, ds1$total_cases_per_million, use="na.or.complete")
#
cor(ds1$icu_patients_per_million, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$hosp_patients_per_million, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$weekly_icu_admissions_per_million, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$weekly_hosp_admissions_per_million, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$hospital_beds_per_thousand, ds1$total_cases_per_million, use="na.or.complete")
#
cor(ds1$population_density, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$median_age, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$aged_65_older, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$aged_70_older, ds1$total_cases_per_million, use="na.or.complete")
#
cor(ds1$cardiovasc_death_rate, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$diabetes_prevalence, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$female_smokers, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$male_smokers, ds1$total_cases_per_million, use="na.or.complete")
cor(ds1$handwashing_facilities, ds1$total_cases_per_million, use="na.or.complete")
```


Slide 31: Decision Tree on New Cases
```{r slide 31, warning=FALSE}
#Prepare data (for new cases rate)
df <- ds %>%
  filter(!grepl('^OWID.*',iso_code))  %>% 
  filter(date >= '2020-03-01' & date <= '2022-03-31')
df <- df %>% 
  select(new_cases_smoothed_per_million, stringency_index, icu_patients_per_million, hosp_patients_per_million, weekly_icu_admissions_per_million,
         weekly_hosp_admissions_per_million, population_density, median_age, aged_65_older, aged_70_older, gdp_per_capita, 
         extreme_poverty, cardiovasc_death_rate, diabetes_prevalence, female_smokers, male_smokers, handwashing_facilities,
         hospital_beds_per_thousand, human_development_index)

factor_mean <- mean(df$new_cases_smoothed_per_million, na.rm=T) + sd(df$new_cases_smoothed_per_million, na.rm=T)
df$is_high <- ifelse(df$new_cases_smoothed_per_million > factor_mean, 1, 0)
df <- df %>% drop_na(is_high)
df$is_high <- factor(df$is_high)
#
df <- df[-1]

set.seed(10)  # fix the randomizer
inTraining <- createDataPartition(df$is_high, p = 0.75, list = FALSE)
# Split data in 75%-25% ratio

training <- df[inTraining,] #75% data for model training  
testing <- df[-inTraining,] #25% for model testing

# nrow(df); nrow(training); nrow(testing)
# Build decision tree (new cases)
fit <- rpart(is_high ~ ., data=training, method="class", 
             parms=list(split="information"))
# summary(fit)

pred <- predict(fit, type="class", newdata=testing)   # Predict Class
ConfusionMatrix <- table(pred, testing$is_high)      # Confusion Matrix Table
ConfusionMatrix
sum(diag(ConfusionMatrix))/sum(ConfusionMatrix)       # calculate the accuracy
# Plot the decision tree
fancyRpartPlot(fit, yesno=2, split.col="black", nn.col="black", 
               caption="", palette="Set2", branch.col="black")
```

<b>Part 3: Prediction on Hong Kong 5th wave</b>

Prepare data
```{r part 3 data prep}
ds_hk <- ds %>%
  filter(iso_code %in% c("HKG"))

ds_hk$date <- as.Date(ds_hk$date, "%Y-%m-%d") # Convert date from string to Date() type

# Parameters test (debug only)
# period <- 90
# forecast <- 30
# end_date <- summarise(ds_hk, date = max(date))
# end_date <- end_date[1,1]
# start_date <- end_date - period
# class(start_date)

# ds2 <-na.omit(ds2)
# View(ds2)
# str(ds2)
```


Slide 32: Prepare data for building regression model using no.2-4 COVID-19 waves in HK
```{r slide 32}
# 4th wave
wave4 <- ds_hk %>%
  select(date, new_cases_smoothed) %>%
  filter(date>="2020-11-15" & date<="2021-02-15")

# 3rd wave
wave3 <- ds_hk %>%
  select(date, new_cases_smoothed) %>%
  filter(date>="2020-07-01" & date<="2020-09-10")

# 2nd wave
wave2 <- ds_hk %>%
  select(date, new_cases_smoothed) %>%
  filter(date>="2020-03-01" & date<="2020-04-30")

g1 <- ggplot(data=wave4) +
  geom_line(aes(x=date, y=new_cases_smoothed)) +
  ggtitle("Wave 4th")
g2 <- ggplot(data=wave3) +
  geom_line(aes(x=date, y=new_cases_smoothed)) +
  ggtitle("Wave 3rd")
g3 <- ggplot(data=wave2) +
  geom_line(aes(x=date, y=new_cases_smoothed)) +
  ggtitle("Wave 2nd")

grid.arrange(g3,g2,g1, ncol=3, nrow=1) # Plot 3 waves graphs to find suitable model
```

Slide 33: \
Prepare regression function \
Define a Gaussian function (of four parameters) \
Ref: https://en.wikipedia.org/wiki/Gaussian_function
```{r slide 33, warning=FALSE}
f <- function(x, theta)  { 
  m <- theta[1]; s <- theta[2]; a <- theta[3]; b <- theta[4];
  a*exp(-0.5*((x-m)/s)^2) + b
}
```


Slide 34: Regression the HK 2nd wave 
```{r Slide 34, warning=FALSE}
wave2$num_days <- round(as.numeric(difftime(wave2$date, "2020-03-01" ,units="days"))) # Change date to number of days

x <- wave2$num_days
y <- wave2$new_cases_smoothed

# Estimate some starting values
m.0 <- x[which.max(y)]; s.0 <- (max(x)-min(x))/4; b.0 <- min(y); a.0 <- (max(y)-min(y))

# Do the fit
fit2 <- nls(y ~ f(x,c(m,s,a,b)), data.frame(x,y), start=list(m=m.0, s=s.0, a=a.0, b=b.0))

# Display the estimated location of the peak and its SE
# summary(fit2)$parameters["m", 1:2]
summary(fit2)

# Graph the estimated curve and real data point
par(mfrow=c(1,1))
plot(c(x,0),c(y,f(coef(fit2)["m"],coef(fit2))), main="COVID19 HK 2nd wave regression", type="n",
     xlab="number of days", ylab="New cases")
curve(f(x, coef(fit2)), add=TRUE, col="Red", lwd=2)
points(x,y, pch=19)
```


Slide 35: Regression 3rd wave
```{r slide 35, warning=FALSE}
wave3$num_days <- round(as.numeric(difftime(wave3$date, "2020-07-01" ,units="days")))

x <- wave3$num_days
y <- wave3$new_cases_smoothed

# Estimate some starting values
m.0 <- x[which.max(y)]; s.0 <- (max(x)-min(x))/4; b.0 <- min(y); a.0 <- (max(y)-min(y))

# Do the fit
fit3 <- nls(y ~ f(x,c(m,s,a,b)), data.frame(x,y), start=list(m=m.0, s=s.0, a=a.0, b=b.0))

# Display the estimated location of the peak and its SE
# summary(fit3)$parameters["m", 1:2]
summary(fit3)

# Graph the estimated curve and real data point
par(mfrow=c(1,1))
plot(c(x,0),c(y,f(coef(fit3)["m"],coef(fit3))), main="COVID19 HK 3rd wave regression", type="n",
     xlab="number of days", ylab="New cases")
curve(f(x, coef(fit3)), add=TRUE, col="Red", lwd=2)
points(x,y, pch=19)
```


Slide 36: Regression 4th wave 
```{r slide 36, warning = FALSE}
wave4$num_days <- round(as.numeric(difftime(wave4$date, "2020-11-15" ,units="days")))

x <- wave4$num_days
y <- wave4$new_cases_smoothed

# Estimate some starting values
m.0 <- x[which.max(y)]; s.0 <- (max(x)-min(x))/4; b.0 <- min(y); a.0 <- (max(y)-min(y))

# Do the fit

fit4 <- nls(y ~ f(x,c(m,s,a,b)), data.frame(x,y), start=list(m=m.0, s=s.0, a=a.0, b=b.0))

# Display the estimated location of the peak and its SE
# summary(fit4)$parameters["m", 1:2]
summary(fit4)
# plot(x, resid(fit4), main="Residuals")

# Graph the estimated curve and real data point
par(mfrow=c(1,1))
plot(c(x,0),c(y,f(coef(fit4)["m"],coef(fit4))), main="COVID19 HK 4th wave regression", type="n",
     xlab="number of days", ylab="New cases")
curve(f(x, coef(fit4)), add=TRUE, col="Red", lwd=2)
points(x,y, pch=19)
```

Slide 37: \
Prediction based on above regression fits \
HK Wave 5 data, choose 2022-1-31 as start date \
Training data end at 2022-3-31, will use latest data for prediction
```{r slide 37, warning = FALSE}
wave5 <- ds_hk %>%
  select(date, new_cases_smoothed) %>%
  filter(date>="2022-01-31" & date<="2022-04-15")

wave5$num_days <- round(as.numeric(difftime(wave5$date, "2022-01-31" ,units="days")))

# Predict use fit of 2nd wave
new_cases2 = predict(fit2, wave5, interval = "prediction")

# Predict use fit of 3rd wave
new_cases3 = predict(fit3, wave5, interval = "prediction")

# Predict use fit of 4th wave
new_cases4 = predict(fit4, wave5, interval = "prediction")

# Plot all the above prediction together and the real data of 5th wave 
plot(wave5$num_days[1:length(new_cases2)], new_cases2, type="l", col="green", xlab="", ylab="")
par(new=TRUE)
plot(wave5$num_days[1:length(new_cases3)], new_cases3, type="l", col="orange", xlab="", ylab="")
par(new=TRUE)
plot(wave5$num_days[1:length(new_cases4)], new_cases4, type="l", col="red", xlab="", ylab="")
par(new=TRUE)
plot(wave5$num_days, wave5$new_cases_smoothed, col="blue", xlab="Number of days", ylab="New cases", main="Regression fit curve to 5th wave from 2-4 waves")
```

<b>*** End ***</b>
